DES Publication Title

### 2018-12-20 papers/methods/misc/author_list.tex
Adding author list

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib,manuscript]{emulateapj}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{enumerate}
\usepackage{xcolor}

%\usepackage{tabu}
%\usepackage{tabularx}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\name}{Steve}
\newcommand{\myemail}{samuelreay@gmail.com}
\newcommand\abs[1]{\left|#1\right|}
\newcommand {\etal} {\emph{~et~al.} }
\definecolor{forestgreen}{rgb}{0.13, 0.75, 0.13}
\newcommand{\green}{\color{forestgreen}}
\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\color{red}}

\newcommand{\kmsmpc}{{\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1}}
\newcommand{\cov}{\mathcal{C}_\eta}
\newcommand{\Z}{\mathcal{Z}}

\defcitealias{Guy2010}{G10}
\defcitealias{Chotard2011}{C11}
\defcitealias{Rubin2015}{R15}	
\defcitealias{Scolnic2016}{SK16}	 
\newcommand{\gten}{\citetalias{Guy2010}}
\newcommand{\steve}{\textit{Steve}}
\newcommand{\celeven}{\citetalias{Chotard2011}}
\newcommand{\rubin}{\citetalias{Rubin2015}}
\newcommand{\sk}{\citetalias{Scolnic2016}}
\newcommand{\angstrom}{\mbox{\normalfont\AA}}

%\usepackage{lineno}
%\linenumbers
\renewcommand{\arraystretch}{1.2}
%\graphicspath{ {./figures/} }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.

\shorttitle{{\steve}}
% Don't change these lines
\begin{document}
\title{{\steve}: A hierarchical Bayesian model for Supernova Cosmology}

@@

% Abstract of the paper
\begin{abstract}
We present a new Bayesian hierarchical model (BHM) named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
%
This advances previous works by including 
an improved treatment of Malmquist bias, 
accounting for additional sources of systematic uncertainty, 
and increasing numerical efficiency. 
%
Given light curve fit parameters, redshifts, and host-galaxy masses, we fit {\steve} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties. Selection effects are characterised using Monte-Carlo simulations.
%
We demonstrate its implementation by fitting realisations of SN~Ia datasets where the SN~Ia model closely follows that used in {\steve}.
%
Next, we validate on more realistic SNANA simulations of SN~Ia samples from the Dark Energy Survey and low-redshift surveys \citep{DESKEY}.
%
These simulated datasets contain more than $60\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
%
This is the most rigorous test of a BHM method applied to SN~Ia cosmology fitting, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model. This $w$-bias is less than $0.03$ on average, less than half the statistical uncertainty on $w$.
%
These simulation test results are a concern for BHM cosmology fitting applications on large upcoming surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
\end{abstract}

\keywords{cosmology: supernovae}

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed type Ia supernovae (SN~Ia) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, chambers2016panstarrs, sako2018sdss}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB. For a recent review, see \citet{Huterer2018}. While these efforts have reduced the uncertainty on the equation-of-state of dark energy ($w$) by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.

In attempts to tease out the nature of dark energy, active and planned surveys are continually growing in size and scale. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has discovered thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will discover tens of thousands of photometrically classified supernovae. Such increased statistical power demands greater fidelity and flexibility in modelling supernovae for cosmological purposes, as we will require reduced systematic uncertainties to fully utilise these increased statistics \citep{Betoule2014, Scolnic2017}.

As such, considerable resources are aimed at developing more sophisticated supernova cosmology analyses. The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific supernova models. First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved for SDSS \citep{Kessler2009}, simulations are a fundamental component of modern supernova cosmology.  \citet{Betoule2014} quantise and correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN~Ia distances as a function of multiple factors to improve bias correction. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, and either use simulation-determined distance-corrections to correct for biases, or attempt to find analytic approximations for effects such as Malmquist bias to model the biases inside the BHM itself.

In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015}. We include additional sources of systematic uncertainty, including an analytic formulation of selection efficiency which incorporates parametric uncertainty. We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.

Section \ref{sec:review} is dedicated to a quick review of supernova cosmology analysis methods, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement. We summarise our methodology in Section \ref{sec:conclusion}.

\section{Review}
\label{sec:review}

Whilst supernova observations take the form of photometric time-series brightness measurements in many bands and a redshift measurement of the supernova (or its assumed host), most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}. The resulting output from the model is, for each supernova, an amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics ($\mathcal{C_\eta}$). As all supernovae are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.

This ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population. Accurately characterising this underlying population, its evolution over redshift, and effects from host-galaxy environment, is one of the challenges of supernova cosmology. Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude of the supernova, traditionally in the Bessell $B$ band), $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space. Specifically, for any given supernova our map between absolute magnitude and apparent magnitude is given by the distance modulus:
\begin{equation}
\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
\end{equation}
where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter. $\Delta M \cdot m$ refers to the host-galaxy mass correlation discussed in Section \ref{sec:hostgal}. The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift. The `$\text{other corrections}$' term often includes bias corrections for traditional $\chi^2$ analyses. Bias corrections can take multiple forms, such as a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.

\subsection{Traditional Cosmology Analyses} \label{sec:traditional}

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted distance modulus. The $\chi^2$ function minimised is
\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1}_{\mu} (\mu_{\rm obs} - \mu_C),
\end{equation}
where $C^{-1}_{\mu}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties for the DES supernova analysis). The predicted $\mu_C$ is defined as
\begin{eqnarray}
\mu_C &=& 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L &=& (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &=& \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{eqnarray}
where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood results in a loss of model flexibility by assuming Gaussian uncertainty. Secondly, the method of creating a covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses information about correlation between sources of uncertainty. For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is ignored when determining population uncertainty using numerical derivatives of population permutations. Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods that can automatically capture correlated uncertainties are preferable. Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A recent solution to this computational cost problem is to bin the supernovae in redshift bins, which produces a matrix of manageable size and can allow fitting without matrix inversion at every step. Whilst binning data results in some loss of information, recent works tested against simulations show that this loss does not result in significant cosmological biases \citep{Scolnic2016, Kessler2017}.

Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting the determined $\mu_{\rm obs}$ from the data, or equivalently, adding a distance bias to the $\mu_C$ prediction. Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift. An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}, which simulates supernovae according to equation \eqref{eq:standard}. Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then subtracted from the observational data. When using traditional fitting methods such as that found in \citet{Betoule2014}, these effects are not built into the likelihood and instead are formed by correcting data. This means that the bias uncertainty is not captured fully in the $\chi^2$ distribution, and subtle correlations between cosmological or population parameters and the bias correction is lost. Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections in the $\chi^2$ likelihood.

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{malmquist.pdf}
	\end{center}
	\caption{An example of the effects of Malmquist bias. Here are shown 1000 simulated supernovae redshifts and distance modulus given fiducial cosmology. The simulated survey is magnitude limited, and all supernovae brighter than magnitude 24 are successfully observed (shown as blue dots), and all dimmer than 24th magnitude are not successfully observed (shown as red dots). By binning the supernovae along redshift, and taking the mean distance modulus of the supernovae in each bin, we can see that at higher redshift where Malmquist bias kicks in, the population mean drops and becomes biased. This source of bias must either be corrected by adjusting the data (such as subtracting the found bias) for by incorporating Malmquist bias explicitly in the cosmological model.}
	\label{fig:malmquist}
\end{figure}

\subsection{Approximate Bayesian Computation}

To avoid the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions. \citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples. In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.

Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects associated with supernova surveys. By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied, and we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood such as those found in traditional analyses from Section \ref{sec:traditional}. This freedom comes with a cost -- computation. The classical $\chi^2$ method's most computationally expensive step in a fit is matrix inversion. For ABC methods, we must instead simulate an entire supernova population -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts. This is an intensive process.

One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$, and $c$). \citet{Jennings2016} presents two metrics, which are used to measure the distance between the forward modelled population and observed population, and are minimised in fitting. The first metric compares forward modelled summary statistic populations (denoted the `Tripp' metric) and the second utilises the observed supernova light curves themselves, moving past summary statistics. However, we note that evaluation of systematic uncertainty was only performed using the Tripp metric.

\subsection{Bayesian Hierarchical Models}

Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM). Hierarchical models utilise multiple layers of connected parameters, with the layers linked via well defined and physically motivated conditional probabilities. For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter. We can thus incorporate different population distributions, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix. Unlike ABC methods, which can model arbitrary probability distributions, BHM methods are generally constrained to representing probabilities using analytic forms.

With the introduction of multiple layers in our model, we can add more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light-curve fitting. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias corrections. \citet{March2014} and \citet{Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS-like simulations. The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset \citep[using redshift dependent bias corrections from][]{Betoule2014}, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}. Notably, these methods rely on data that is bias corrected or the methods ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The assumption made by the UNITY analysis is that the bias in real data is perfectly described by an analytic function. They validate their  model to be free of significant biases using fits to thirty realisations of supernova datasets that are constructed to mimic the UNITY framework. The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, most lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}. For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.

Whilst there are a large number of hierarchical models available, none of them have undergone comprehensive tests using realistic simulations to verify each models' respective bias. Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to both $\Omega_m$ and $w$.

The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, color and stretch populations, host-galaxy corrections, and redshift evolution, each of which will be discussed further in the outline of our model below. Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models. We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections in an effort to provide a valuable cross-check on analysis methodologies which utilise scatter-model-dependant bias corrections.

\section{Challenges in Supernova Cosmology}
\label{sec:challenges}

 The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by analyses. In this Section, we outline several of the most prominent challenges.

 Forefront among these challenges is our ignorance of the underlying type Ia intrinsic dispersion. Ideally, analysis of the underlying dispersion would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae. However this data is difficult to obtain, and recent efforts to quantify the dispersion draw inference from photometric measurements. The underlying dispersion model is not a solved problem, and we therefore test against two dispersion models in this work. The first is based on the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, the second is sourced from \citet[][hereafter denoted {\celeven}]{Chotard2011}. As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C_\eta}$. These two scatter models are currently assumed to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.

The underlying supernova population is further complicated by the presence of outliers. Non-type Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae  and represent outliers from the standardised SN~Ia sample. This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts  being assigned to supernovae. Different optimizations to the host-galaxy algorithm can result in misidentification of the host at the 3\% to 9\% level \citep{Gupta2016}, resulting in a broad population of outliers. For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible. However, for photometric surveys, which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit, and correct for contaminants.

Finally, one of the other persistent challenges facing supernova cosmology analyses are the high number of systematics. Because of the rarity of SN~Ia events, datasets are commonly formed from the SN~Ia discoveries of multiple surveys in order to increase the number of supernovae in a dataset. However, each survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys. Peculiar velocities, different host environments, and dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.

\section{Our Method}
\label{sec:method}

We construct our hierarchical Bayesian model {\steve} with several goals in mind: creation of a redshift-dependent underlying supernova population,  treatment of an increased number of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections. We also desire {\steve} more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days. As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in the programming language Stan \citep{Carpenter2017, StanDevelopmentTeam2017}. The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS, a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.

At the most fundamental level, a supernova cosmology analysis is a mapping from an underlying population onto an observed population, where cosmological parameters are encoded directly in the mapping function. The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both the observed and underlying populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model. In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies. Summaries of observables and model parameters are shown in Table \ref{tab:param_summary} for easy reference.

\subsection{Observed Populations}

\subsubsection{Observables}
Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a brightness measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and color $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$. Additionally, each supernova has an observed redshift $\hat{z}$ and a host-galaxy stellar mass associated with it, $\hat{m}$, where the mass measurement is converted into a probability of being above $10^{10}$ solar masses. We also have a probability of each supernova being a type Ia, $\hat{p}$. Our set of observables input into the {\steve} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement {\blue as varying redshifts introduce computational complexity in computing the distance modulus integral}) but instead uncertainty on distance modulus. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate. One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, these misclassifications can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.

\begin{deluxetable}{ll}
	\tablecolumns{2}
	\tablewidth{0.99\columnwidth}
	\tablecaption{Model Parameters}
	\tablehead { \colhead{Parameter} & \colhead{Description} }
	\startdata

	&\textbf{Global Parameters} \\

	$\Omega_m$  & Matter density  \\
	$w$  &  Dark energy equation of state  \\
	$\alpha$ & Stretch standardisation   \\
	$\beta$         &  color standardisation   \\
	$\delta(0)$ & Scale of the mass-magnitude correction\\
	$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
	$\delta\mathcal{Z}_i$ & Systematics scale\\
	$\langle M_B \rangle$ & Mean absolute magnitude \\
	\\
	& \textbf{Survey Parameters} \\

	$\delta S$ & Selection effect deviation \\
	$\langle x_1^i \rangle$ & Mean stretch nodes\\
	$\langle c^i \rangle$ & Mean color nodes\\
	$\alpha_c$ & Skewness of color population \\
	$\sigma_{M_B}$ & Population magnitude scatter \\
	$\sigma_{x_1}$ & Population stretch scatter \\
	$\sigma_{c}$ & Population color scatter \\
	$\kappa_{0}$ & Extra color dispersion\\
	$\kappa_{1}$ & Redshift-dependence of extra color dispersion\\
	\\
	& \textbf{Supernova Parameters} \\
	$m_B$ & True flux\\
	$x_1$ & True stretch \\
	$c$ & True color \\
	$z$ & True redshift \\	
	$M_B$ & Derived absolute magnitude \\
	$\mu$ & Derived distance modulus \\
	\\
	& \textbf{Input Data}\tablenotemark{a} \\

	$\hat{m}_B$ & Measured flux\\
	$\hat{x}_1$ & Measured stretch \\
	$\hat{c}$ & Measured color \\
	$C$ & Covariance on flux, stretch and color \\
	$\hat{z}$ & Observed redshift \\
	$\hat{m}$ & Determined mass probability \\
	$\hat{p}$ & Determined Ia probability \\
	\enddata
	\tablenotetext{a}{Not model parameters but shown for completeness.}
\label{tab:param_summary}
\end{deluxetable}

\subsubsection{Latent Variables for Observables}

The first layer of hierarchy is the set of true (latent) parameters that describe each supernova. In contrast to the observed parameters, the latent parameters are denoted without a hat. For example, $c$ is the true color of the supernova, whilst $\hat{c}$ is the observed color, which, as it has measurement error, is different from $c$.

For the moment, let us consider a single supernova and its classic summary statistics $m_B$, $x_1$, $c$. For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$. A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves fluxes and uncertainties. However, this is computationally prohibitive  as it would require incorporating SALT2 light curve fitting inside our model fitting. Due to this computational expense, we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C_\eta}$ describing the uncertainty on $\hat{\eta}$. Using this simplification, our latent variables are given by
\begin{equation}
p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C_\eta}). \label{eq:pop}
\end{equation}
As discussed in Section \ref{sec:challenges},  the SALT2 model does not include full treatment of intrinsic dispersion, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.

 \begin{figure}
 	\begin{center}
 		\includegraphics[width=\columnwidth]{tikz-figure0.pdf}
 	\end{center}
 	\caption{Probabilistic graphical model for our likelihood. Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables. The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it. Variables that appear outside both boxes represent top level model parameters.}
 	\label{fig:pgm}
 \end{figure}

\subsection{Underlying Population}

\subsubsection{Type Ia population} 
\label{sec:underlying}

Unlike many previous formalisms which utilise $M_B$ as a singular number and model magnitude scatter on the apparent magnitude $m_B$, we incorporate this scatter into the underlying rest-frame population by having a population in absolute magnitude space. {\blue This is mathematically equivalent, however allows us to model the underlying population and intrinsic scatter distinctly.} To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.

In addition to absolute magnitude, the underlying supernova population is also characterised by distributions in color and stretch. We follow the prior work of {\rubin} and model the color population as an independent redshift-dependent skew normal distribution for each survey. For the stretch population, we adopt a redshift-dependent normal distribution, and magnitude dispersion is modelled as a normal distribution. We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no reduction in cosmological bias with the skew-normal distributions for stretch and magnitude. Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the color and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component) to capture correlations between $m_B$ and $c$, however pathological fitting complications required us to simplify our treatment. We parametrise the color skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ that allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}. We sample $\delta_c$ in log space for efficiency in sampling close to zero. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors with {\blue mean 0 and width 1}.

The only constant between survey populations is the absolute magnitude $\langle M_B \rangle$. We model the colour skewness and the redshift-dependent means and width of the colour and skew distributions individually for each survey. The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
\begin{eqnarray}
P(M_B, x_1, c, z | \theta) &=& \mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \, \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1})\notag\\ &\quad\quad& \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1},
\end{eqnarray}
where $\theta = \lbrace \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c \rbrace$ for legibility.

\subsubsection{Outlier populations}

For future use with photometrically classified surveys, we include a simplistic outlier population model. We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model. For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SN~Ia population in prior works. However, with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.  We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.  We note that this represents the mean brightness of outliers, and so outliers could both be brighter and dimmer than the mean SN~Ia absolute magnitude. We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$ discussed in the previous section. For the spectroscopic survey, we set this to unity. For the photometric proof-of-concept we provide an accurate probability estimate. Further investigation on the effect of inaccurate estimates will be left for future improvements during the analysis of the DES photometric sample.

\subsection{Correcting biased summary statistics}

With the fitted summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology. We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model. We should also avoid assuming a specific dispersion model -- either the {\gten} or {\celeven} model, or utilise the results of computing the bias from both models.

>>>>

% Author list file generated with: mkauthlist.py 1.2.4 
% mkauthlist.py -f --doc -j emulateapj --sort --sort-builder -a order.csv author_list.csv author_list.tex

\documentclass[iop]{emulateapj}

\begin{document}
\title{DES Publication Title}

\def\andname{}

@@

\begin{abstract}
This is a sample document created by \texttt{mkauthlist.py v1.2.4}.
\end{abstract}
\maketitle
\end{document}
